---
output:
  html_document:
    keep_md: yes
---
LOTR R Markdown file
========================================================

This R Markdown file presents the original data, the statistical summaries and all figures generated.

# Data Tidying and Preliminary Analysis

Load libraries

```{r}
library(ggplot2)
library(dplyr)
library(stringr)
```

Import data

```{r}
lotr_data <- read.delim("lotr_raw.tsv")
glimpse(lotr_data)
```

Make Chapter column into two columns (number and name) and rename columns after binding together

```{r}
lotr_data_split <- data.frame(do.call(rbind, strsplit(as.vector(lotr_data$Chapter), split = ":")))
names(lotr_data_split) <- c("Chapter_Number", "Chapter_Name")
lotr_data_clean_split <- cbind(lotr_data$Film, lotr_data_split, lotr_data$Character, lotr_data$Race, lotr_data$Words)
ldcs2<- plyr::rename(lotr_data_clean_split, c("lotr_data$Film"="Film", "lotr_data$Character"="Character", "lotr_data$Race"="Race", "lotr_data$Words"="Words"))
ldcs2
```

Reorder Film factor levels

```{r}
old_levels <- levels(ldcs2$Film)
j_order <- sapply(c("Fellowship", "Towers", "Return"),
				  function(x) grep(x, old_levels))
new_levels <- old_levels[j_order]
```

Clean up dataset: apply new factor levels to Film, change name of races, drop less frequent races, arrange based on race,film, words

```{r}
lotr_data_clean <- ldcs2 %>% 
	mutate("Film" = factor(as.character(Film), new_levels),
		   "Race" = plyr::revalue(Race, c(Ainur="Wizard", Men="Man"))) %>% 
	filter(!(Race %in% c("Gollum", "Orc", "Ent", "Dead", "Nazgul"))) %>% 
	mutate(Race = reorder(Race, Words, sum)) %>% 
	arrange(Race,Film, Character, Words) %>% 
	droplevels
glimpse(lotr_data_clean)
```

Here is a plot about which race does the most talking by film

```{r}
plot1 <- lotr_data_clean %>% 
	ggplot(aes(Race, Words)) + 
	geom_col() +
	facet_wrap(~Film) +
	labs(y="Total Words Spoken") 
print(plot1)
```

the pre-made plot from above that I am embedding in the compiled document.


# The Nine Companions

Filter to show total words spoken by the Nine Companions

```{r}
fellowship <- c("Gandalf", "Aragorn", "Merry", "Pippin", "Frodo", "Sam", "Gimli", "Legolas", "Boromir")
(characterwords <- lotr_data %>% 
	filter(Character %in% fellowship) %>% 
	group_by(Character, Film) %>% 
	summarize(sum(Words)))
```

For each Film, show a plot of words by fellowship member

```{r}
plyr::d_ply(characterwords, ~Film, function(p){
	the_film <- p$Film[1]
	R <- ggplot(p, aes(Character, `sum(Words)`))+ 
		geom_col() +
		theme(axis.text.x=element_text (angle=270)) +
		labs(y="Total Words Spoken", x= "Nine Companions")+
		ggtitle(the_film)
	print(R)
})
```


